Fusing Filters with Integer Linear Programming

### 2014-04-15 2014betterfusionforfilters/section/03-SizeInference.tex
attempt to turn size inference into natural deduction proofs. ugly. why can't I center them?

\begin{tabbing}
MMMMM       \= MM \= MMMMMM \= \kill
$rate$      \> @::=@ \> $Filtered_n$        \> (output of @filter@s) \\
            \> $~|$  \> $External_n$        \> (output of @external@s) \\
            \> $~|$  \> $rate \times rate$  \> (cross product) \\
            \> $~|$  \> $k_n$               \> (rate variable)\\
\end{tabbing}

@@

\begin{tabbing}
MMMMM       \= MM \= MMMM \= MM \= \kill
$constrs$  \> @::@  \> $bind$  \> $\to$ \> $\{var = rate\}$ \\
MMMMMMMMMMMMMM               \= MM  \= \kill
$constrs(o = @fold@~f~n)  $ \> $=$ \> $\{\}$     \\
$constrs(o = @map@_n~f~ns)$ \> $=$ \> $\{o = n~|~n \in ns\}$     \\
$constrs(o = @filter@~f~b)$ \> $=$ \> $\{o = Filtered_o\}$              \\
$constrs(o = @gather@~i~d)$ \> $=$ \> $\{o = i\}$ \\
$constrs(o = @cross@~a~b) $ \> $=$ \> $\{o = a \times b\}$ \\
$constrs(outs = @external@~ins) $ \> $=$ \> $\{o = External_o~|~o \in outs\}$ \\
\end{tabbing}

Then, for some program $\lambda\ldots~@let@~binds~@in@~\ldots$, find the union of all constraints:
$\bigcup_{b \in binds} constrs(b)$.
The constraints are all equalities, so the generated equalities are easily grouped into equivalence classes.

All combinators whose constraints are in the same equivalence class have the same iteration size, and may be fused together.
These constraints are also used to rule out any programs that require runtime checks on array lengths during the program.

>>>>

\begin{tabbing}
MMMMM       \= MM \= MMMMMM \= \kill
$rate$      \> @::=@ \> $k_n$               \> (rate variable)\\
            \> $~|$  \> $rate \times rate$  \> (cross product) \\
%            \> $~|$  \> $Filtered_n$        \> (output of @filter@s) \\
%            \> $~|$  \> $External_n$        \> (output of @external@s) \\
\\
$constraint$ \> @::=@ \> $rate = rate$                      \\
             \> $~|$  \> $rate \le rate$                    \\
             \> $~|$  \> $constraint \wedge constraint$     \\
             \> $~|$  \> $\forall k_n. constraint$          \\
             \> $~|$  \> $\exists k_n. constraint$          \\
\end{tabbing}

@@

\vspace{1em}

\infer{\lambda scalars~arrays \to @let @ binds : \forall_{a \in arrays} k_a. C}
      { \{k_a | a \in arrays\} \vdash binds : C}

\vspace{1em}

\infer{\Gamma \vdash o = @fold@~f~n; binds : C}
      {\Gamma \vdash binds                 : C}

\vspace{1em}

\infer{\Gamma \vdash o = @map@_n~f~ns; binds : \forall k_o. \bigwedge_{n \in ns}\{k_o = k_n\} \wedge C}
      {k_o, \Gamma \vdash binds                 : C}

\vspace{1em}

\infer{\Gamma \vdash o = @filter@~f~n; binds : \exists k_o. k_o \le k_n \wedge C}
      {k_o, \Gamma \vdash binds                 : C}

\vspace{1em}

\infer{\Gamma \vdash o = @gather@~i~d; binds : \forall k_o. k_o = k_i \wedge C}
      {k_o, \Gamma \vdash binds                 : C}

\vspace{1em}

\infer{\Gamma \vdash o = @cross@~a~b; binds : \forall k_o. k_o = k_a \times k_b \wedge C}
      {k_o, \Gamma \vdash binds                 : C}

\vspace{1em}

\infer{\Gamma \vdash outs = @external@~ins; binds : \exists_{o \in outs} k_o. C}
      {\{k_o | o \in outs\} \cup \Gamma \vdash binds                 : C}

\vspace{1em}

The generated equalities are easily grouped into equivalence classes.
All combinators whose constraints are in the same equivalence class have the same iteration size, and may be fused together.
These constraints are also used to rule out any programs that require runtime checks on array lengths during the program.
